---
title: "Graphing with Scaled Data"
author: "Calvin Skalla"
date: "4/22/2020"
output: html_document
---

# Loading packages and importing/cleaning the data
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(GGally)

#Reading in the Data
white <- read.csv("data/winequality-white.csv", sep = ";")
red <- read.csv("data/winequality-red.csv", sep = ";")

#Combining and Clarifying Type
wine_data <- rbind(white, red) %>% tibble::rowid_to_column("ID") %>% mutate(Type = ifelse(ID <= 4898, "white", "red")) %>% select(-ID)

#Making quality a factor for easier visualization
wine_data$quality <- as.factor(wine_data$quality)

#normalizing data
preproc2 <- preProcess(wine_data, method = c("range"))
w_data_minmax <- predict(preproc2, wine_data)

#Subsetting dataset for easier visualizations
twentypercent <- sample_frac(w_data_minmax, size = 0.20)
```

#Scatterplot Matrices for Correlations
```{r, echo=FALSE, warning=FALSE, message=FALSE}
cor(x = w_data_minmax[,1:11], y = as.numeric(w_data_minmax$quality))
#Scatterplot Matrix demonstrating correlation between predictor variables
ggpairs(data = w_data_minmax, columns = c("volatile.acidity", "citric.acid", "chlorides", "free.sulfur.dioxide", "density", "alcohol"), lower = list(continuous = wrap("points", size = 1)))
```

# Chemical Distributions for each Quality
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Violin Plots showing the distributions for each variable over the qualities
fixed.acidity <- ggplot(twentypercent, aes(x = quality, y = fixed.acidity)) + geom_violin() + coord_flip()

volatile.acidity <- ggplot(twentypercent, aes(x = quality, y = volatile.acidity)) + geom_violin() + coord_flip()
volatile.acidity

citric.acid <- ggplot(twentypercent, aes(x = quality, y = citric.acid)) + geom_violin() + coord_flip()
citric.acid

residual.sugar <- ggplot(twentypercent, aes(x = quality, y = residual.sugar)) + geom_violin() + coord_flip()

chlorides <- ggplot(twentypercent, aes(x = quality, y = chlorides)) + geom_violin() + coord_flip()
chlorides

free.sulfur.dioxide <- ggplot(twentypercent, aes(x = quality, y = free.sulfur.dioxide)) + geom_violin() + coord_flip()

total.sulfur.dioxide <- ggplot(twentypercent, aes(x = quality, y = total.sulfur.dioxide)) + geom_violin() + coord_flip()

density <- ggplot(twentypercent, aes(x = quality, y = density)) + geom_violin() + coord_flip()
density

pH <- ggplot(twentypercent, aes(x = quality, y = pH)) + geom_violin() + coord_flip()

sulphates <- ggplot(twentypercent, aes(x = quality, y = sulphates)) + geom_violin() + coord_flip()

#Effective
alcohol <- ggplot(twentypercent, aes(x = quality, y = alcohol)) + geom_violin() + coord_flip()
alcohol
```

# Coplots
```{r, echo=FALSE, warning=FALSE, message=FALSE}
density_alcohol <- ggplot(w_data_minmax, aes(alcohol, density)) +
    geom_point(size = 0.5, color = "lightgrey") +
    facet_wrap(~cut_width(quality, width = 2, center = 5)) +
    geom_smooth()

density_alcohol
```

# Density plots of wine quality, Red vs. White
```{r}
library(ggridges)
library(patchwork)
p1 <- ggplot(wine_data) + geom_bar(
  aes(x = quality, y = ..prop.., fill = Type, group = Type), 
  position = position_dodge(),
  stat = 'count')
p2 <- ggplot(wine_data) + geom_density_ridges(
  aes(x = quality, y = Type, height = ..density.., group = Type, fill = Type),
  alpha = 0.7)
p1 / p2
```

# Comparison of alcohol percentage by type

```{r}
wine_data %>% ggplot(aes(x = Type, y = alcohol, fill = Type)) +
  geom_boxplot(alpha = 0.4) +
  geom_violin(fill = NA, scale = "width", linetype = 2)
```

```{r}
melted <- reshape2::melt(w_data_minmax, id.vars = c('quality', 'Type'), na.rm = TRUE)
melted %>% group_by(quality, Type, variable) %>%
  summarise(mean = mean(as.numeric(value))) %>%
  ungroup() %>%
  arrange(desc(mean)) %>%
  ggplot() + geom_col(aes(x = quality, y = mean, fill = variable), 
                      stat = 'identity',
                      color = 'gray',
                      position = position_dodge()) +
  coord_flip() +
  facet_wrap(~ Type)
```

```{r}
melted %>% filter((Type == 'white' & quality == 9) | (Type == 'red' & quality == 8)) %>%
  group_by(Type, quality, variable) %>%
  summarise(mean = mean(as.numeric(value))) %>%
  ungroup() %>%
  arrange(desc(quality)) %>%
  ggplot() + geom_col(aes(x = variable, y = mean, fill = variable), 
                      stat = 'identity',
                      color = 'gray',
                      position = position_dodge()) +
  coord_polar() +
  facet_wrap(~Type) +
  ggtitle("Feature Breakdown of Top Rated Reds and Whites")
```

# Classification Tree

```{r}
library(caret)
library(rpart)
library(e1071)

# Changing all character features to factors
wine_data.factors <- wine_data %>%
  mutate_if(is.character, as.factor)

# Set up our CV training methods
train_control <- trainControl(method = "cv", 
                              number = 10)

# Create our train/test split.
set.seed(42)
train_index <- sample(1:nrow(wine_data.factors), 0.9 * nrow(wine_data))
test_index <- setdiff(1:nrow(wine_data.factors), train_index)

train_dat <- wine_data.factors[train_index, ]
test_dat <- wine_data.factors[test_index, ]

# Growing our first tree
tree.model <- train(quality ~ ., data = train_dat, 
                    method = "rpart", 
                    trControl = train_control, 
                    na.action = na.rpart)
```

```{r}
library(rpart.plot)
rpart.plot(tree.model$finalModel, yesno = 2)
rpart.rules(tree.model$finalModel, cover = TRUE)
conmat <- confusionMatrix(data = tree.model)
print(conmat)
```

Single tree doesn't yield very high accuracy, but the single tree is nice for illustration. Let's try a random forest.

```{r}
library(randomForest)
forest.model <- randomForest(quality ~ ., data = train_dat)
```

```{r}
imp <- varImp(forest.model, scale = TRUE)
imp <- data.frame(rownames(imp), imp$Overall)
names(imp) <- c('Factor', 'Importance')

imp <- imp %>% transform(Factor = reorder(Factor, Importance))
ggplot(imp, aes(Factor, Importance)) + 
  geom_bar(stat = 'identity', color = "blue", fill = "white") + 
  coord_flip() +
  ggtitle("Variable Importance in Determining Wine Quality")
```

```{r}
(conm <- forest.model$confusion)
```

Better than the single tree! Let's visualize...

```{r}
library(viridis)
# Drop the last column to make things easier
conm.v <- conm[, -8]
conm.v.perc <- (conm.v / apply(conm.v, 1, sum)) * 100
melted <- reshape2::melt(conm.v.perc)

ggplot(melted, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = viridis(10)[1],
    mid = viridis(10)[5],
    high = viridis(10)[10],
    midpoint = 50,
    limit = c(0, 100),
    name = "Accuracy Percentage"
  ) +
  geom_text(aes(label = round(value, 1)), color = "white") +
  xlab("Predicted Quality") + ylab("True Quality") + 
  ggtitle("Confusion Matrix for Random Forest Predictions") +
  scale_x_continuous(breaks = seq(3, 9), labels = seq(3, 9), expand = c(0,0)) +
  scale_y_continuous(breaks = seq(3, 9), labels = seq(3, 9), expand = c(0,0))
```

And try it out on our holdout test data...

```{r}
preds <- predict(forest.model, test_dat)
(conm <- confusionMatrix(preds, test_dat$quality))
```

```{r}
conm.v <- conm$table
conm.v.perc <- (conm.v / apply(conm.v, 1, sum)) * 100
melted <- reshape2::melt(conm.v.perc)

ggplot(melted, aes(x = Prediction, y = Reference, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = viridis(10)[1],
    mid = viridis(10)[5],
    high = viridis(10)[10],
    midpoint = 50,
    limit = c(0, 100),
    name = "Accuracy Percentage"
  ) +
  geom_text(aes(label = round(value, 1)), color = "white", na.rm = TRUE) +
  xlab("Predicted Quality") + ylab("True Quality") + 
  ggtitle("Confusion Matrix for Random Forest Predictions") +
  scale_x_continuous(breaks = seq(3, 9), labels = seq(3, 9), expand = c(0,0)) +
  scale_y_continuous(breaks = seq(3, 9), labels = seq(3, 9), expand = c(0,0))
```

Overall, not bad! Around 70% overall accuracy, and fairly high overall focused on the diagonal, and relatively low (or 0) percentage predictions further out from the diagonal.